Pillar 1 - Student Topic Profile

### 2019-06-07 Warning Issuance - Grade Prediction with RF.Rmd
explore randomForest package

```{r find_df} 
find_df <- function(course)  d_student_model %>% filter(`Course ID` == course)
```

Consider courses currently on offer and with at least 20 enrollments since 2008 
```{r}
d_lasso_course <- tibble(target = course_current) %>%

@@

# Random forest

```{r}
data(iris)

@@

```{r}
rf1 <- randomForest(Species ~ ., iris, ntree=50, norm.votes=FALSE)

@@

rf.all <- combine(rf1, rf2, rf3)
print(rf.all)
```

```{r}
## Look at the third trees in the forest.
getTree(randomForest(iris[,-5], iris[,5], ntree=10), 3, labelVar=TRUE)
```

```{r}
iris.rf <- randomForest(Species ~ ., iris, ntree=50, norm.votes=FALSE)
iris.rf <- grow(iris.rf, 50)
print(iris.rf)
```

```{r}
set.seed(4543)
data(mtcars)
mtcars.rf <- randomForest(mpg ~ ., data=mtcars, ntree=1000,
keep.forest=FALSE, importance=TRUE)

@@

```{r}
data(imports85)
imp85 <- imports85[,-2] # Too many NAs in normalizedLosses.
imp85 <- imp85[complete.cases(imp85), ]
## Drop empty levels for factors.
imp85[] <- lapply(imp85, function(x) if (is.factor(x)) x[, drop=TRUE] else x)
stopifnot(require(randomForest))
price.rf <- randomForest(price ~ ., imp85, do.trace=10, ntree=100)
print(price.rf)
numDoors.rf <- randomForest(numOfDoors ~ ., imp85, do.trace=10, ntree=100)
print(numDoors.rf)
```

>>>>

Consider courses currently on offer and with at least 20 enrollments since 2008 
```{r}
d_course <- tibble(target = course_current) %>%

@@

# Random Forest

```{r}
## Regression:
data(airquality)
ozone.rf <- randomForest(Ozone ~ ., data=airquality, mtry=3, ntree = 1e3,
                         importance=TRUE, na.action=na.omit)
print(ozone.rf)
## Show "importance" of variables: higher value mean more important:
round(importance(ozone.rf), 2)
varImpPlot(ozone.rf)

# margin
plot(randomForest::margin(iris.rf), F)
plot(randomForest::margin(iris.rf), T)

# MDSplot
MDSplot(iris.rf, iris$Species)

# Looping over variables ranked by importance:
airquality <- na.omit(airquality)
ozone.rf <- randomForest(Ozone ~ ., airquality, importance=TRUE)
imp <- importance(ozone.rf)
impvar <- rownames(imp)[order(imp[, 1], decreasing=TRUE)]
op <- par(mfrow=c(2, 3))

for (i in seq_along(impvar))
  partialPlot(ozone.rf, 
              airquality,
              impvar[i], 
              xlab=impvar[i], main=paste("Partial Dependence on", impvar[i]), ylim=c(30, 70))
par(op)

# predict

ind <- sample(2, nrow(iris), replace = TRUE, prob=c(0.8, 0.2))
iris.rf <- randomForest(Species ~ ., data=iris[ind == 1,])

iris.pred <- predict(iris.rf, iris[ind == 2,])
table(observed = iris[ind==2, "Species"], predicted = iris.pred)

## Get prediction for all trees.
predict(iris.rf, iris[ind == 2,], predict.all=TRUE)

# Tune randomForest for the optimal mtry parameter
data(fgl, package="MASS")
fgl.res <- tuneRF(fgl[,-10], fgl[,10], ntreeTry = 1000, stepFactor=1.5)

# CV for for feature selection
myiris <- cbind(iris[1:4], matrix(runif(96 * nrow(iris)), nrow(iris), 96))
result <- rfcv(myiris, iris$Species, cv.fold=3)
with(result, plot(n.var, error.cv, log="x", type="o", lwd=2))

## The following can take a while to run, so if you really want to try
## it, copy and paste the code into R.

## Not run: 
result <- replicate(5, rfcv(myiris, iris$Species), simplify=FALSE)
error.cv <- sapply(result, "[[", "error.cv")
matplot(result[[1]]$n.var, cbind(rowMeans(error.cv), error.cv), type="l",
        lwd=c(2, rep(1, ncol(error.cv))), col=1, lty=1, log="x",
        xlab="Number of variables", ylab="CV Error")

## End(Not run)

# tree size
hist(treesize(iris.rf))
```

TODO: combine RF of all courses to evaluate feature importance

# Random forest trial
```{r randomForest}
## Classification:
data(iris)
iris.rf <- randomForest(Species ~ ., data=iris, importance=TRUE,
                        proximity=TRUE)
print(iris.rf)
## Look at variable importance:
round(importance(iris.rf), 2)
## Do MDS on 1 - proximity:
iris.mds <- cmdscale(1 - iris.rf$proximity, eig=TRUE)
op <- par(pty="s")
pairs(cbind(iris[,1:4], iris.mds$points), cex=0.6, gap=0,
      col=c("red", "green", "blue")[as.numeric(iris$Species)],
      main="Iris Data: Predictors and MDS of Proximity Based on RandomForest")
par(op)
print(iris.mds$GOF)

## The `unsupervised' case:
set.seed(17)
iris.urf <- randomForest(iris[, -5])
MDSplot(iris.urf, iris$Species)

## stratified sampling: draw 20, 30, and 20 of the species to grow each tree.
(iris.rf2 <- randomForest(iris[1:4], iris$Species, 
                          sampsize=c(20, 30, 20)))

## Regression:
data(airquality)
ozone.rf <- randomForest(Ozone ~ ., data=airquality, mtry=3,
                         importance=TRUE, na.action=na.omit)
print(ozone.rf)
## Show "importance" of variables: higher value mean more important:
round(importance(ozone.rf), 2)

## "x" can be a matrix instead of a data frame:
set.seed(17)
x <- matrix(runif(5e2), 100)
y <- gl(2, 50)
(myrf <- randomForest(x, y))
(predict(myrf, x))

## "complicated" formula:
(swiss.rf <- randomForest(sqrt(Fertility) ~ . - Catholic + I(Catholic < 50),
                          data=swiss))
(predict(swiss.rf, swiss))
## Test use of 32-level factor as a predictor:
set.seed(1)
x <- data.frame(x1=gl(53, 10), x2=runif(530), y=rnorm(530))
(rf1 <- randomForest(x[-3], x[[3]], ntree=10))

## Grow no more than 4 nodes per tree:
(treesize(randomForest(Species ~ ., data=iris, maxnodes=4, ntree=30)))

## test proximity in regression
iris.rrf <- randomForest(iris[-1], iris[[1]], ntree=101, proximity=TRUE, oob.prox=FALSE)
str(iris.rrf$proximity)
```

```{r classCenter}
data(iris)

@@

```{r randomForest::combine}
rf1 <- randomForest(Species ~ ., iris, ntree=50, norm.votes=FALSE)

@@

rf.all <- randomForest::combine(rf1, rf2, rf3)
print(rf.all)
```

```{r getTree}
## Look at the third trees in the forest.
getTree(randomForest(iris[,-5], iris[,5], ntree=10), k = 3, labelVar=TRUE)
```

```{r grow}
iris.rf <- randomForest(Species ~ ., iris, ntree=50, norm.votes=FALSE)
iris.rf <- grow(iris.rf, 50)
print(iris.rf)
```

```{r importance}
set.seed(4543)
data(mtcars)
mtcars.rf <- randomForest(mpg ~ ., data=mtcars, ntree=1000,
                          keep.forest=FALSE, importance=TRUE)

@@

```{r imports85}
data(imports85)
imp85 <- imports85[,-2] # Too many NAs in normalizedLosses.
imp85 <- imp85[complete.cases(imp85), ]
## Drop empty levels for factors.
imp85[] <- lapply(imp85, function(x) if (is.factor(x)) x[, drop=TRUE] else x)
stopifnot(require(randomForest))

price.rf <- randomForest(price ~ ., imp85, do.trace=10, ntree=100) # regression
print(price.rf)

numDoors.rf <- randomForest(numOfDoors ~ ., imp85, do.trace=10, ntree=100) # classification
print(numDoors.rf)
```

```{r margin}
iris.rf <- randomForest(Species ~ ., iris, keep.forest=FALSE)
plot(randomForest::margin(iris.rf), F)
plot(randomForest::margin(iris.rf), T)
```

```{r MDSplot}
iris.rf <- randomForest(Species ~ ., iris, proximity=TRUE,
                        keep.forest=FALSE)
MDSplot(iris.rf, iris$Species)
## Using different symbols for the classes:
MDSplot(iris.rf, iris$Species, palette=rep(1, 3), pch=as.numeric(iris$Species))
```

```{r na.roughfix}
## artificially drop some data values.
iris.na <- iris
for (i in 1:4) iris.na[sample(150, sample(20)), i] <- NA

iris.roughfix <- na.roughfix(iris.na)
iris.narf <- randomForest(Species ~ ., iris.na, na.action=na.roughfix)
print(iris.narf)
```

```{r outlier}
iris.rf <- randomForest(iris[,-5], iris[,5], proximity=TRUE)
plot(outlier(iris.rf), type="h",
     col=c("red", "green", "blue")[as.numeric(iris$Species)])
```

```{r partialPlot}
iris.rf <- randomForest(Species~., iris)
partialPlot(iris.rf, iris, Petal.Width, "versicolor")

## Looping over variables ranked by importance:
airquality <- na.omit(airquality)
ozone.rf <- randomForest(Ozone ~ ., airquality, importance=TRUE)
imp <- importance(ozone.rf)
impvar <- rownames(imp)[order(imp[, 1], decreasing=TRUE)]
op <- par(mfrow=c(2, 3))

for (i in seq_along(impvar))
  partialPlot(ozone.rf, 
              airquality,
              impvar[i], 
              xlab=impvar[i], main=paste("Partial Dependence on", impvar[i]), ylim=c(30, 70))
par(op)
```

```{r plot.randomForest}
data(mtcars)
plot(randomForest(mpg ~ ., mtcars, keep.forest=FALSE, ntree=100), log="y")
```

```{r predict.randomForest}
ind <- sample(2, nrow(iris), replace = TRUE, prob=c(0.8, 0.2))
iris.rf <- randomForest(Species ~ ., data=iris[ind == 1,])

iris.pred <- predict(iris.rf, iris[ind == 2,])
table(observed = iris[ind==2, "Species"], predicted = iris.pred)

## Get prediction for all trees.
predict(iris.rf, iris[ind == 2,], predict.all=TRUE)

## Proximities.
predict(iris.rf, iris[ind == 2,], proximity=TRUE)

## Nodes matrix.
str(attr(predict(iris.rf, iris[ind == 2,], nodes=TRUE), "nodes"))
```

```{r rfcv}
myiris <- cbind(iris[1:4], matrix(runif(96 * nrow(iris)), nrow(iris), 96))
result <- rfcv(myiris, iris$Species, cv.fold=3)
with(result, plot(n.var, error.cv, log="x", type="o", lwd=2))

## The following can take a while to run, so if you really want to try
## it, copy and paste the code into R.

## Not run: 
result <- replicate(5, rfcv(myiris, iris$Species), simplify=FALSE)
error.cv <- sapply(result, "[[", "error.cv")
matplot(result[[1]]$n.var, cbind(rowMeans(error.cv), error.cv), type="l",
        lwd=c(2, rep(1, ncol(error.cv))), col=1, lty=1, log="x",
        xlab="Number of variables", ylab="CV Error")

## End(Not run)
```

```{r rfImpute}
iris.na <- iris
set.seed(111)
## artificially drop some data values.
for (i in 1:4) iris.na[sample(150, sample(20)), i] <- NA

set.seed(222)
iris.imputed <- rfImpute(Species ~ ., iris.na)

set.seed(333)
iris.rf <- randomForest(Species ~ ., iris.imputed)

print(iris.rf)
```

```{r treesize}
iris.rf <- randomForest(Species ~ ., iris)
hist(treesize(iris.rf))
```

```{r tuneRF}
data(fgl, package="MASS")
fgl.res <- tuneRF(fgl[,-10], fgl[,10], ntreeTry = 1000, stepFactor=1.5)
```

```{r varImpPlot}
mtcars.rf <- randomForest(mpg ~ ., data=mtcars, ntree=1000, keep.forest=FALSE,
                          importance=TRUE)
varImpPlot(mtcars.rf)
```

```{r varUsed}
varUsed(randomForest(Species~., iris, ntree=100))
```
